Evaluating FAIR Digital Object as a distributed object system

### 2023-02-16 latex/main.tex
Updates from Overleaf

\documentclass[fleqn,10pt,lineno]{wlpeerjlua}

@@

% https://www.overleaf.com/latex/examples/bibliographies-with-biber-and-biblatex/ccrkczqwnywf
% use biblatex/biber as natbib is from 1990s and have no DOIs
\usepackage[autostyle]{csquotes}
\usepackage[backend=biber,alldates=iso,labeldate=year,style=authoryear]{biblatex}
\addbibresource{bibliography.bib}
\DeclareFieldFormat{doi}{%
  \url{https://doi.org/#1}}
%\urlstyle{tt}

@@

\hypertarget{sec:introduction}{%
\section*{Introduction}\label{sec:introduction}}

@@

\hypertarget{sec:background}{%
\section*{Background and related work}\label{sec:background}}

@@

\hypertarget{sec:fdo}{%
\subsection*{FAIR Digital Object}\label{sec:fdo}}

@@

\hypertarget{fdo-approaches}{%
\subsubsection*{FDO approaches}\label{fdo-approaches}}

@@

\hypertarget{sec:next-step-fdo}{%
\subsubsection*{Next steps for FDO}\label{sec:next-step-fdo}}

@@

\hypertarget{sec:ld}{%
\subsection*{From the Semantic Web to Linked Data}\label{sec:ld}}

@@

\hypertarget{sec:semweb}{%
\subsubsection*{A brief history of the Semantic Web}\label{sec:semweb}}

@@

\hypertarget{sec:ld-web}{%
\subsubsection*{Linked Data: Rebuilding the Web of Data}\label{sec:ld-web}}

@@

\hypertarget{sec:comparing}{%
\subsection*{Comparing FDO and existing approaches}\label{sec:comparing}}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  An Interoperability Framework and Distributed Platform for Fast Data Applications \autocite{delgadoInteroperabilityFrameworkDistributed2016a}, which proposes quality measurements for comparing how frameworks support interoperability, particularly from a service architectural view
\item
  The FAIR Digital Object guidelines \autocite{boninoFAIRDigitalObject}, validated against its implementations for completeness.
\item
  A Comparison Framework for Middleware Infrastructures \autocite{zarrasComparisonFrameworkMiddleware2004a}, which suggest dimensions like openness, performance and transparency, mainly focused on remote computational methods
\item
  Cross-checks against RDA's FAIR Data Maturity Model \autocite{bahimFAIRDataMaturity2020a} to find how the FAIR principles are achieved in FDO, in particular considering access, sharing and openness
\item
  EOSC Interoperability Framework \autocite{corchoEOSCInteroperabilityFramework2021b} which gives recommendations for technical, semantic, organisational and legal interoperability, particularly from a metadata perspective
\end{enumerate}

@@

\section*{Methods}

\hypertarget{sec:interoperability-compare}{%
\subsection*{Considering FDO/Web as interoperability framework for Fast Data}\label{sec:interoperability-compare}}

@@

A difference between Web technologies and FDO is the stringency of the requirements for both syntax and semantics. Whereas the Web allows many different syntactic formats (e.g.~from HTML to XML, PDFs), FDO realised with DOIP requires JSON. On the semantic front, FDO requires that every object have a well-defined type and structured form. This is clearly not the case on the Web.

In terms of connectivity and the deployment of applications, the Web has a plethora of software, services, and protocols that are widely deployed. These have shown interoperability. The Web standards bodies (e.g.~IETF and W3C) follow the OpenStand principles \autocite{ModernStandardsParadigm} to embrace openness, transparency, and broad consensus. In contrast, FDO has a small number of implementations and corresponding protocols, although with a growing community, as evidenced at the first FDO conference \autocite{looFirstInternationalConference2022}. This is not to say that it is not worth developing further Handle+DOIP implementations in the future, but we note that the current FDO functionality can easily be implemented using Web technologies, even as DOIP-over-HTTP \autocite{DOIPAPIHTTPa}.

@@

\hypertarget{mapping-of-metamodel-concepts}{%
\subsubsection*{Mapping of Metamodel concepts}\label{mapping-of-metamodel-concepts}}

@@

\hypertarget{sec:doip-fdo-compare}{%
\subsection*{Assessing FDO implementations}\label{sec:doip-fdo-compare}}

@@

\hypertarget{sec:middleware}{%
\subsection*{Comparing FDO and Web as middleware infrastructures}\label{sec:middleware}}

@@

\hypertarget{sec:fair-compare}{%
\subsection*{Assessing FDO against FAIR}\label{sec:fair-compare}}

In addition to having ``FAIR'' in its name, the FAIR Digital Object guidelines \autocite{fdo-RequirementSpec} also include \emph{G3: FDOs must offer compliance with the FAIR principles through measurable indicators of FAIRness.} .

@@

\hypertarget{eosc-interoperability-framework}{%
\section*{EOSC Interoperability Framework}\label{eosc-interoperability-framework}}

@@

\hypertarget{sec:discussion}{%
\section*{Discussion}\label{sec:discussion}}

@@

\begin{itemize}
\tightlist
\item
  Ramnifications of ide
\item
  Finish tables
\item
  Bullet points per table
\item
  Send in Google Docs
\end{itemize}

\hypertarget{what-does-it-mean-for-linked-data}{%
\subsection*{(What does it mean for Linked Data?)}\label{what-does-it-mean-for-linked-data}}

@@

>>>>

\documentclass[fleqn,10pt,NOlineno]{wlpeerjlua}

@@

% https://www.overleaf.com/latex/examples/bibliographies-with-biber-and-biblatex/ccrkczqwnywf
% use biblatex/biber as natbib is from 1990s and have no DOIs
\usepackage[autostyle]{csquotes}
\usepackage[backend=biber,alldates=iso,labeldate=year,style=authoryear]{biblatex}
\addbibresource{bibliography.bib}
\DeclareFieldFormat{doi}{%
  \url{https://doi.org/#1}}
\usepackage{biblatex2bibitem}  
%\urlstyle{tt}

@@

\section*{Introduction}\label{sec:introduction}

@@

\section*{Background and related work}\label{sec:background}

@@

\subsection*{FAIR Digital Object}\label{sec:fdo}

@@

\subsubsection*{FDO approaches}\label{sec:fdo-approaches}

@@

\subsubsection*{Next steps for FDO}\label{sec:next-step-fdo}

@@

\subsection*{From the Semantic Web to Linked Data}\label{sec:ld}

@@

\subsubsection*{A brief history of the Semantic Web}\label{sec:semweb}

@@

\subsubsection*{Linked Data: Rebuilding the Web of Data}\label{sec:ld-web}

@@

\subsection*{Comparing FDO and existing approaches}\label{sec:comparing}

@@

\begin{enumerate}
\tightlist
\item
  An Interoperability Framework and Distributed Platform for Fast Data Applications \autocite{delgadoInteroperabilityFrameworkDistributed2016a}, which proposes quality measurements for comparing how frameworks support interoperability, particularly from a service architectural view
\item
  The FAIR Digital Object guidelines \autocite{boninoFAIRDigitalObject}, validated against its implementations for completeness.
\item
  A Comparison Framework for Middleware Infrastructures \autocite{zarrasComparisonFrameworkMiddleware2004a}, which suggest dimensions like openness, performance and transparency, mainly focused on remote computational methods
\item
  Cross-checks against RDA's FAIR Data Maturity Model \autocite{bahimFAIRDataMaturity2020a} to find how the FAIR principles are achieved in FDO, in particular considering access, sharing and openness
\item
  EOSC Interoperability Framework \autocite{corchoEOSCInteroperabilityFramework2021b} which gives recommendations for technical, semantic, organisational and legal interoperability, particularly from a metadata perspective
\end{enumerate}

@@

\section*{Methods}
\subsection*{Considering FDO/Web as interoperability framework for Fast Data}\label{sec:interoperability-compare}

@@

A difference between Web technologies and FDO is the stringency of the requirements for both syntax and semantics. Whereas the Web allows many different syntactic formats (e.g.~from HTML to XML, PDFs), FDO realised with DOIP requires JSON. On the semantic front, FDO requires that every object have a well-defined type and structured form. This is clearly not the case on the Web.
In terms of connectivity and the deployment of applications, the Web has a plethora of software, services, and protocols that are widely deployed. These have shown interoperability. The Web standards bodies (e.g.~IETF and W3C) follow the OpenStand principles \autocite{ModernStandardsParadigm} to embrace openness, transparency, and broad consensus. In contrast, FDO has a small number of implementations and corresponding protocols, although with a growing community, as evidenced at the first FDO conference \autocite{looFirstInternationalConference2022}. This is not to say that it is not worth developing further Handle+DOIP implementations in the future, but we note that the current FDO functionality can easily be implemented using Web technologies, even as DOIP-over-HTTP \autocite{DOIPAPIHTTPa}.

@@

\subsubsection*{Mapping of Metamodel concepts}\label{mapping-of-metamodel-concepts}

@@

\subsection*{Assessing FDO implementations}\label{sec:doip-fdo-compare}

@@

\subsection*{Comparing FDO and Web as middleware infrastructures}\label{sec:middleware}

@@

\subsection*{Assessing FDO against FAIR}\label{sec:fair-compare}

In addition to having ``FAIR'' in its name, the FAIR Digital Object guidelines \autocite{fdo-RequirementSpec} also include \emph{G3: FDOs must offer compliance with the FAIR principles through measurable indicators of FAIRness}.

@@

\section*{EOSC Interoperability Framework}\label{eosc-interoperability-framework}

@@

\section*{Discussion}\label{sec:discussion}

@@

\begin{itemize}
\tightlist
\item
  Ramnifications of ide
\item
  Bullet points per table
\end{itemize}

\subsection*{(What does it mean for Linked Data?)}\label{what-does-it-mean-for-linked-data}

@@

% To generate bibitems for manual latex editing, enable
%\printbibitembibliography
